---
title: "Gap Statistic"
author: "Group 01"
date: "14 maggio 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(plotly)
library(mvtnorm)
```

# Simulations for the Gap Statistic

We're going to start by defining a function which, given a matrix $X$ containing the points we are interested in clustering, gives a sample from the reference distribution associated to $X$.

```{r cars}
reference_sample = function(points_matrix, sample_size, method = "svd") {
    if(method == "svd") {
        return(reference_SVD_sample(points_matrix, sample_size))
    } else if(method == "simple") {
        return(reference_simple_sample(points_matrix, sample_size))
    } else {
        stop("Not valid method for reference sampling!")
    }
}

reference_simple_sample = function(points_matrix, sample_size) {
    mins = c()
    maxs = c()
    for (column in 1 : ncol(points_matrix)) {
        mins = c(mins, min(points_matrix[, column]))
        maxs = c(maxs, max(points_matrix[, column]))
    }
    samples = c()
    for (column in 1 : ncol(points_matrix)) {
        samples = c(samples, runif(sample_size, min = mins[column], max = maxs[column]))
    }
    return(matrix(samples, ncol = ncol(points_matrix), byrow = FALSE))
}

reference_SVD_sample = function(points_matrix, sample_size) {
    means = c()
    for (column in 1 : ncol(points_matrix)) {
        means = c(means, mean(points_matrix[, column]))
    }
    centered_matrix = points_matrix - means
    V = svd(centered_matrix)$v
    sample = reference_simple_sample(centered_matrix %*% V, sample_size)
    sample = sample %*% t(V)
    return(sample + means)
}
```

## Including Plots

We now define a function which, given a set of points, computes the gap statistic for that set under the kmean clustering algorithm.

```{r pressure, echo=FALSE}
compute_gap_statistic = function(data, k, B, method = "svd", with_se = FALSE) {
    standard_data = scale(data)
    clustering = kmeans(data, k, iter.max = 25)
    log_WK = log(clustering$tot.withinss)
    rows = nrow(data)
    if(length(c(B)) == 1) {
        reference_sample_data = reference_sample(data, B * rows, method)
        groups = B
    } else {
        reference_sample_data = B
        groups = nrow(reference_sample_data) / rows
    }
    log_WKStar = c()
    for (instance in 1 : groups) {
        first = 1 + rows * (instance - 1)
        last = rows * instance
        clustering = kmeans(reference_sample_data[first : last, ], k, iter.max = 25)
        log_WKStar = c(log_WKStar, log(clustering$tot.withinss))
    }
    if(with_se) {
        se = sd(log_WKStar) * sqrt(1 + (1 / groups))
        return(c(mean(log_WKStar) - log_WK, se))
    }
    return(mean(log_WKStar) - log_WK)
}
```

And then we code a function to find out the best $K$, number of clusters, for a set of points.

```{r pressure, echo=FALSE}
find_best_k = function(data, B, method = "svd") {
    gap = -1
    k = 0
    reference_sample_data = reference_sample(data, B * nrow(data), method)
    gap_k_plus = compute_gap_statistic(data, k + 1, reference_sample_data, method)
    while(gap < 0) {
        k = k + 1
        gap_k = gap_k_plus
        k_plus = compute_gap_statistic(data, k + 1, reference_sample_data, method, TRUE)
        gap_k_plus = k_plus[1]
        se_k_plus = k_plus[2]
        gap = gap_k - gap_k_plus + se_k_plus
    }
    return(k)
}
```

## Simulations

### Resampling methods

First of all we want to show how the two different resampling methods work. We'll see that the SVD method is better when there's a good amount of correlation in the data, as it can retrievs the variance structure of the distribution.

```{r pressure, echo=FALSE}
data_figure = function(data, trace_name) {
    return(plot_ly(x = data[, 1], y = data[, 2], type = "scatter", mode = "markers", name = trace_name))
}

simulation_data = matrix(rmvnorm(100, c(0, 0), diag(c(1, 1))), nrow = 100, byrow = TRUE)
simple_sample = reference_sample(simulation_data, 100, method = "simple")
svd_sample = reference_sample(simulation_data, 100, method = "svd")

simulation_figure = data_figure(simulation_data, "Original Data")
simple_figure = data_figure(simple_sample, "Simple Resampling")
svd_sample = data_figure(svd_sample, "SVD Resampling")

figure = subplot(simulation_figure, simple_figure, svd_sample, shareY = TRUE)
figure = figure %>% layout(title = "Normal Standard Distribution")
figure
```



```{r}
simulation_data = c()
for (observation in 1 : 100) {
    x = rnorm(1, 0, 2)
    y = rnorm(1, x, 1)
    simulation_data = c(simulation_data, x, y)
}
simulation_data = matrix(simulation_data, nrow = 100, byrow = TRUE)
simple_sample = reference_sample(simulation_data, 100, method = "simple")
svd_sample = reference_sample(simulation_data, 100, method = "svd")

simulation_figure = data_figure(simulation_data, "Original Data")
simple_figure = data_figure(simple_sample, "Simple Resampling")
svd_sample = data_figure(svd_sample, "SVD Resampling")

figure = subplot(simulation_figure, simple_figure, svd_sample, shareY = TRUE)
figure = figure %>% layout(title = "High Correlated Data")
figure
```

### Example of application of the Gap Statistic

```{r pressure, echo=FALSE}
simulation_data = c()

for (observations in 1 : 30) {
    simulation_data = c(simulation_data, rmvnorm(1, c(-1, -1), diag(c(0.5, 0.5))))
    simulation_data = c(simulation_data, rmvnorm(1, c(3, 1.5), diag(c(1, 2))))
    simulation_data = c(simulation_data, rmvnorm(1, c(3, 1.5), diag(c(1, 2))))
}
simulation_data = matrix(simulation_data, nrow = 90, byrow = TRUE)
colors = rep(c("red", "green", "green"), 30)

simulation_data_figure = plot_ly(x = simulation_data[, 1], y = simulation_data[, 2], color = colors, type = "scatter", mode = "markers", name = "Data")
simulation_data_figure = simulation_data_figure %>% layout(plot_bgcolor = 'rgb(236, 236, 243)')

clusters = 1 : 10
Wk = c()
for (K in clusters) {
    clustering = kmeans(simulation_data, K)
    Wk = c(Wk, clustering$tot.withinss)
}

Wk_figure = plot_ly(x = clusters, y = Wk, type = "scatter", mode = "lines+markers", name = "Wk", height = 600)
Wk_figure = Wk_figure %>% layout(yaxis = list(title = "Wk"))

gaps = c()
sds = c()
for (K in clusters) {
    couple = compute_gap_statistic(simulation_data, K, 1000, with_se = TRUE)
    gaps = c(gaps, couple[1])
    sds = c(sds, couple[2])
}

best_k = -1
for (K in clusters) {
    if(gaps[K] > gaps[K + 1] - sds[K + 1]) {
        best_k = K
        break
    }
}

best_k_line = list(type = "line", y0 = 0, y1 = 1, yref = "paper", x0 = best_k, x1 = best_k, line = list(dash = "dot", color = "purple", thickness = 0.5))
treshold_line = list(type = "line", y0 = gaps[best_k + 1] - sds[best_k + 1], y1 = gaps[best_k + 1] - sds[best_k + 1], x0 = best_k + 1, x1 = best_k + 0.2, line = list(dash = "dot", color = "purple", thickness = 0.5))

bands = list(array = sds, thickness = 1, color = "blue")
gap_figure = plot_ly(x = clusters, y = gaps, type = "scatter", mode = "lines+markers", name = "Gap(k)", error_y = bands)
gap_figure = gap_figure %>% layout(shapes = list(best_k_line, treshold_line))
Wk_figure = Wk_figure %>% layout(shapes = list(best_k_line))

figure = subplot(simulation_data_figure, subplot(Wk_figure, gap_figure, nrows = 2))
figure
```

```{r pressure, echo=FALSE}
simulation_data = c()

for (observations in 1 : 30) {
    simulation_data = c(simulation_data, rmvnorm(1, c(-1, -1), diag(c(0.5, 0.5))))
    simulation_data = c(simulation_data, rmvnorm(1, c(3, 1.5), diag(c(1, 2))))
    simulation_data = c(simulation_data, rmvnorm(1, c(3, 1.5), diag(c(1, 2))))
}
simulation_data = matrix(simulation_data, nrow = 90, byrow = TRUE)
colors = rep(c("red", "green", "green"), 30)

simulation_data_figure = plot_ly(x = simulation_data[, 1], y = simulation_data[, 2], color = colors, type = "scatter", mode = "markers", name = "Data")
simulation_data_figure = simulation_data_figure %>% layout(plot_bgcolor = 'rgb(236, 236, 243)')

clusters = 1 : 10
Wk = c()
for (K in clusters) {
    clustering = kmeans(simulation_data, K)
    Wk = c(Wk, clustering$tot.withinss)
}

Wk_figure = plot_ly(x = clusters, y = Wk, type = "scatter", mode = "lines+markers", name = "Wk", height = 600)
Wk_figure = Wk_figure %>% layout(yaxis = list(title = "Wk"))

gaps = c()
sds = c()
for (K in clusters) {
    couple = compute_gap_statistic(simulation_data, K, 1000, with_se = TRUE)
    gaps = c(gaps, couple[1])
    sds = c(sds, couple[2])
}

best_k = -1
for (K in clusters) {
    if(gaps[K] > gaps[K + 1] - sds[K + 1]) {
        best_k = K
        break
    }
}

best_k_line = list(type = "line", y0 = 0, y1 = 1, yref = "paper", x0 = best_k, x1 = best_k, line = list(dash = "dot", color = "purple", thickness = 0.5))
treshold_line = list(type = "line", y0 = gaps[best_k + 1] - sds[best_k + 1], y1 = gaps[best_k + 1] - sds[best_k + 1], x0 = best_k + 1, x1 = best_k + 0.2, line = list(dash = "dot", color = "purple", thickness = 0.5))

bands = list(array = sds, thickness = 1, color = "blue")
gap_figure = plot_ly(x = clusters, y = gaps, type = "scatter", mode = "lines+markers", name = "Gap(k)", error_y = bands)
gap_figure = gap_figure %>% layout(shapes = list(best_k_line, treshold_line))
Wk_figure = Wk_figure %>% layout(shapes = list(best_k_line))

figure = subplot(simulation_data_figure, subplot(Wk_figure, gap_figure, nrows = 2))
figure
```

```{r}
simulation_data = c()

for (observations in 1 : 40) {
    simulation_data = c(simulation_data, rmvnorm(1, c(-1, -1), diag(c(0.5, 0.5))))
    simulation_data = c(simulation_data, rmvnorm(1, c(3, 1.5), diag(c(1, 2))))
    simulation_data = c(simulation_data, rmvnorm(1, c(3, 1.5), diag(c(1, 2))))
    simulation_data = c(simulation_data, rmvnorm(1, c(-1, 7), diag(c(0.01, 0.02))))
}
simulation_data = matrix(simulation_data, nrow = 160, byrow = TRUE)
colors = rep(c("red", "green", "green", "blue"), 40)

simulation_data_figure = plot_ly(x = simulation_data[, 1], y = simulation_data[, 2], color = colors, type = "scatter", mode = "markers", name = "Data")
simulation_data_figure = simulation_data_figure %>% layout(plot_bgcolor = 'rgb(236, 236, 243)')

clusters = 1 : 10
Wk = c()
for (K in clusters) {
    clustering = kmeans(simulation_data, K)
    Wk = c(Wk, clustering$tot.withinss)
}

Wk_figure = plot_ly(x = clusters, y = Wk, type = "scatter", mode = "lines+markers", name = "Wk", height = 600, line = list(color = "rgb(65, 188, 65)"), marker = list(color = "rgb(65, 188, 65)"))
Wk_figure = Wk_figure %>% layout(yaxis = list(title = "Wk"))

gaps = c()
sds = c()
for (K in clusters) {
    couple = compute_gap_statistic(simulation_data, K, 1000, with_se = TRUE)
    gaps = c(gaps, couple[1])
    sds = c(sds, couple[2])
}

best_k = -1
for (K in clusters) {
    if(gaps[K] > gaps[K + 1] - sds[K + 1]) {
        best_k = K
        break
    }
}

best_k_line = list(type = "line", y0 = 0, y1 = 1, yref = "paper", x0 = best_k, x1 = best_k, line = list(dash = "dot", color = "purple", thickness = 0.5))
treshold_line = list(type = "line", y0 = gaps[best_k + 1] - sds[best_k + 1], y1 = gaps[best_k + 1] - sds[best_k + 1], x0 = best_k + 1, x1 = best_k + 0.2, line = list(dash = "dot", color = "purple", thickness = 0.5))

bands = list(array = sds, thickness = 1, color = "blue")
gap_figure = plot_ly(x = clusters, y = gaps, type = "scatter", mode = "lines+markers", name = "Gap(k)", error_y = bands, line = list(color = "rgb(215, 49, 51)"), marker = list(color = "rgb(215, 49, 51)"))
gap_figure = gap_figure %>% layout(shapes = list(best_k_line, treshold_line))
Wk_figure = Wk_figure %>% layout(shapes = list(best_k_line))

figure = subplot(simulation_data_figure, subplot(Wk_figure, gap_figure, nrows = 2))
figure
```

```{r}
simulation_data = c()

for (observations in 1 : 20) {
    X0 = rnorm(1, -2, 0.3)
    X1 = rnorm(1, -2, 0.3)
    X2 = rnorm(1,  0, 0.2)
    X3 = rnorm(1,  0, 0.2)
    X4 = rnorm(1,  0, 0.2)
    X5 = rnorm(1, 1, 0.01)
    simulation_data = c(simulation_data, X0, X0, X1, X1, X2, X2, X3, X3, X4, X4, X5, X5)
}
simulation_data = matrix(simulation_data, nrow = 120, byrow = TRUE)
colors = rep(c("red", "red", "green", "green", "green", "blue"), 20)

simulation_data_figure = plot_ly(x = simulation_data[, 1], y = simulation_data[, 2], color = colors, type = "scatter", mode = "markers", name = "Data")
simulation_data_figure = simulation_data_figure %>% layout(plot_bgcolor = 'rgb(236, 236, 243)')

clusters = 1 : 10
Wk = c()
for (K in clusters) {
    clustering = kmeans(simulation_data, K)
    Wk = c(Wk, clustering$tot.withinss)
}

Wk_figure = plot_ly(x = clusters, y = Wk, type = "scatter", mode = "lines+markers", name = "Wk", height = 600, line = list(color = "rgb(65, 188, 65)"), marker = list(color = "rgb(65, 188, 65)"))
Wk_figure = Wk_figure %>% layout(yaxis = list(title = "Wk"))

gaps = c()
sds = c()
for (K in clusters) {
    couple = compute_gap_statistic(simulation_data, K, 1000, with_se = TRUE)
    gaps = c(gaps, couple[1])
    sds = c(sds, couple[2])
}

best_k = -1
for (K in clusters) {
    if(gaps[K] > gaps[K + 1] - sds[K + 1]) {
        best_k = K
        break
    }
}

best_k_line = list(type = "line", y0 = 0, y1 = 1, yref = "paper", x0 = best_k, x1 = best_k, line = list(dash = "dot", color = "purple", thickness = 0.5))
treshold_line = list(type = "line", y0 = gaps[best_k + 1] - sds[best_k + 1], y1 = gaps[best_k + 1] - sds[best_k + 1], x0 = best_k + 1, x1 = best_k + 0.2, line = list(dash = "dot", color = "purple", thickness = 0.5))

bands = list(array = sds, thickness = 1, color = "blue")
gap_figure = plot_ly(x = clusters, y = gaps, type = "scatter", mode = "lines+markers", name = "Gap(k)", error_y = bands, line = list(color = "rgb(215, 49, 51)"), marker = list(color = "rgb(215, 49, 51)"))
gap_figure = gap_figure %>% layout(shapes = list(best_k_line, treshold_line))
Wk_figure = Wk_figure %>% layout(shapes = list(best_k_line))

figure = subplot(simulation_data_figure, subplot(Wk_figure, gap_figure, nrows = 2))
figure
```